#Results

```{r source R, warning=F, message=F, echo=FALSE}
## REMOVE CHUNK LATER
source('~/thesis/rcode/plot_ideas.R', echo=FALSE)
```

Some statistically significant results are found from my regression.

```{r echo=F}
lm_senate_D <- sum_cong_lm(senate_D, 67, 113)
lm_senate_R <- sum_cong_lm(senate_R, 67, 113)
lm_house_D <- sum_cong_lm(house_D, 67, 113)
lm_house_R <- sum_cong_lm(house_R, 67, 113)
```

\begin{table}[h]
\centering
\caption{Summary of Models Across All Years, House and Senate}
\label(FullModels)
\begin{tabular}{r|rrrr}
                       & Senate Ds & Senate Rs & House Ds & House Rs \\ \hline
(Intercept)            & 0.00489   & 0.00259   & 0.01131  & 0.00405  \\
cong\_ideo\_mean       & 0.51583   & 0.61344   & 0.73047  & 0.67749  \\
last\_cong\_ideo\_mean & -0.38112  & -0.50989  & -0.57483 & -0.53415 \\
last\_ideo             & 0.87467   & 0.87061   & 0.87636  & 0.85013  \\
leaderY                & \textbf{-0.01863}  & 0.01289   & -\textbf{0.03022} & -0.00417 \\
Adjusted R-Squared     & 0.802     & 0.781     & 0.781    & 0.824    \\
N                      & 2239      & 1817      & 11380    & 9270    
\end{tabular}
\end{table}


All of the control variables that were enlisted were found to be statistically signicant. These three variables were the idelogical mean of congress, the ideological mean of congress from the previous congressional session, and the congressperson's ideology from the previous congressional session. For all three of these variables across both parties in both the House and Senate the p-value was always less than 0.0001. Please see Table 3.1 to see

.

.

.

.

.

.

There are five different outcomes that are possible with the above methodology.

  1. A null result is found - the leaders and non-leaders do not change in significantly different ways.
  2. Leaders move to be more conservative, or right-leaning, when compared to non-leaders.
  3. Leaders move to be more liberal, or left-leaning, when compared to non-leaders.
  4. Leaders move to be more moderate, or centrist, when compared to non-leaders.
  5. Leaders move to be more extreme when compared to non-leaders.
  
There are compelling reasons for many of these to be true. For outcome where the null result is found this could make sense if leaders and non-leaders have little reason to be different. One could imagine leaders becoming more conservative compared to non-leaders due to age of many of the leaders within Congress and the relation between ideology and age. They could become more moderate in order to fascilitate the dealmaking that must be done in Congress in order to get legislation passed. Or perhaps they would becoem more extreme in order to placate their more extreme bases, only choosing to delineate from them in the rarest of circumstances.