# Dataset and Methodology

## Dataset

The dataset that I am using a special version of the DW-NOMINATE dataset that was created by Timothy Nokken and Keith Poole for a 2004 study to research congressional party defection throughout the history of the United States. What it shares with the DW-NOMINATE dataset is a first and second dimasion coordinate for each legislator across the Congresses. But it differs in that

In order to properly subset my data I completely removed all independents from the dataset. In a later study it might be more useful to instead mark independents as a member of the Republican or Democratic party if they decide to caucus with them, such as Senator Sanders, an independent who caucuses with the Democrats, or Senator Joe Lieberman, and independent who also caucused with the Democrats. This did not seem to be a huge reduction in the total dataset size, as during the period this study looks at there were only sixteen total third party members of the Senate and only thirty-two total third party members in the House.\footnote{A significant percentage of these thirty-two are members of the Minnesota Farmer-Labor party or members of the Progressive party.} The variables a reader of this paper might be most interested in are:

- Congress Number, 1st Congress through 113th Congress
- ICPSR ID Number: A unique five digit code assigned to each legislator.
- State Code: A two digit code assigned by state.
- State Name: The name of the state the legislator is from.
- Party Code: 100 if a Democrat, 200 if a Republican, and many others exist for third parties.
- Name: The legislator's last name and occasionally first name and middle name initials.
- 1st Dimension Coordinate: The most important dimension, can place a legislator in a party with over 80% accuracy across most Congresses, and with 100% accuracy in the current polarized environment.
- 2nd Dimension Coordinate: The second dimension of voting data pulled out from the roll call votes. Across history it has often reflected disagreement within the parties. For example, in the 1960s within the Democratic and Republican party it reflected views on Civil Rights. Currently the dimension is not very predictive, but does help designate who is a member of the "establishment" versus the anti-establishment Members of Congress.

New variables were added during the cleaning and coding of the data. They are:

- Role: Blank if none, "W" if whip, "L" if leader, and "S" if Speaker.
- Leadership Position: "Y" if Role is not blank, "N" if Role is blank.
- Party Mean: The mean 1st dimension score of the party for the current Congress. Continuous and Numerical.
- Past Party Mean: The mean 1st dimension score of the party for the past Congress. Continuous and Numerical.
- Past Ideology: The 1st Dimension score for the Member of Congress for the past Congress.

The mean ideologies by party for each chamber are given below to offer greater context of how parties move within their chamber and in particular the extreme polarization of the past few years.


```{r senate_means_table, results = "asis", echo=F}
senate_means_short_10 <- senate_means_short %>% filter(cong > 103)
kable(senate_means_short_10, col.names = c("Congress", "Dem. Ideology", "Rep. Ideology"),
      caption = "Senate Mean Ideology By Party, 104th to 113th Congress")
```

```{r house_means_table, results = "asis", echo=F}
house_means_short_10 <- house_means_short %>% filter(cong > 103)
kable(house_means_short_10, col.names = c("Congress", "Dem. Ideology", "Rep. Ideology"),
      caption = "House Mean Ideology By Party, 104th to 113th Congress")
```

Here you can see a 

## Methodology

My goal is to understand how leadership positions affect the ideology of a Member of Congress. A Member of Congress during one Congress is my unit of observation. I looked at four independent variables and one dependent variable. The independent variables are:

- Past ideology: The 1st dimension score of the Member of Congress for the past Congress. Continuous and Numerical.
- Past party ideology: The mean 1st dimension score of the party for the past Congress. Continuous and Numerical.
- Current party ideology: The mean 1st dimension score of the party for the current Congress. Continuous and Numerical.
- Leadership status: If the Member of Congress is a "leader," as described above in the dataset section. The variable of note. Binary.

I am looking to eventually use these variables to predict the response variable, current ideology. Current ideology is the 1st dimension score of the legislator of interest for the current Congress. I would expect the legislator's past ideology to be the most predictive, followed by the party's current ideology, with the third most predictive varible being the party's past ideology. Leadership status is expected to be the least predictive variable of the four tested for here.



The few assumptions needed for a linear regression model had to be checked for. There is a chance that the model would be better represented by a non-linear model but due to the author's statistical knowhow a linear regression seemed most appropriate. Normality in the error terms as well as a constanct variance of the error terms also seemed to be met. The needed assumption that may have been hardest to work around was multicollinearity, or when predictor terms within the regression are able to predict each other with great accuracy. The variables around ideology could have some issues here and a future paper might try to further tease out whether there is a strong problem of colinearity in this dataset and model.